---
title: "R Notebook"
output: html_notebook
---

## 1

### a
For each sex, split the 10 people randomly into two groups of 5: one for the treadmill and one for the rowing machine. Each "group" of people will have 5 subjects. There will be 2 factors: sex and exercise machine. Sex is observational while the exercise machine is experimental.
### b
Have each person participate in both exercise methods, but randomly assign for each person which machine they exercise first on. There are still two factors.
### c
Sex goes with the larger units while exercise machine goes with the smaller units.

## 2

### a
The server is neither a block nor a unit. Sex itself is observational, not randomized, so using a completely randomized block design is incorrect.
### b
Sex has 1 df, differences within the server has 1 df, the interactions have 1 df, and the error has 96 df.
### c
There is only one server for each block, meaning that there is no replication for Sex. Thus, the experimental design in essence is flawed.
### d
For each server, record tips for 50 consecutive tables, but randomly, for 25 of them, draw a happy face on the back of the check.

## 3

### a
Factor - Sex (Observational, 2 levels: Male, Female), Happy Face (Experimental, 2 levels: Yes or No)
```{r}
setwd("C:/Users/aidan/IdeaProjects/DataScience/data")
knitr::opts_knit$set(root.dir = "C:/Users/aidan/IdeaProjects/DataScience/data")
data.df <- data.frame(Tip=c(0.18, 0.21, 0.33, 0.28), Sex=c("M", "M", "F", "F"), Happy = c("Y", "N", "Y", "N"))
attach(data.df)
interaction.plot(Sex, Happy, Tip,
                 type = "b", pch = c(1,4), col = c("red", "blue"),)
detach()
```

Because the lines are not parallel, there is an interaction between them. There are differences in response from one sex to the another due to whether a happy face was on the back of a check.

## 4

### a

```{r}
oil.df <- read.csv("OilDeapsorbtion.csv", header = T, stringsAsFactors = T)
attach(oil.df)
par(mfrow=c(1,2))
plot(diff ~ Ultra)
plot(diff ~ Oil)
detach()
```

Initially, it seems that an increase in ultrasound may lower the difference in oil deapsorbed. This seems to be the case in the opposite direction for the amount of oil. I think that oil will definitely be significant, however ultrasound not as much.

### b

```{r}
attach(oil.df)
interaction.plot(Ultra, Oil, diff,
                 type = "b", pch = c(1, 4), col = c("red", "blue"))
detach()
```

As the two slopes are not parallel, there is some interaction.

### c

```{r}
attach(oil.df)
par(mfrow = c(2, 2))
library(lattice)
tab <- tapply(diff, list(Ultra, Oil), mean)
adjTab <- data.frame(diff = c(0.5, 0.8, 1.55, 1.10), Ultra = c(5, 10, 5, 10), Oil = c(5, 5, 10, 10))
xyplot(adjTab$diff~adjTab$Ultra,groups=adjTab$Oil,auto.key=list(space="right"),
       panel=function(x,y,...){
         panel.xyplot(x,y,...)
         panel.abline(h=mean(adjTab$diff),col="grey",lty=2,...)
       }
)
plot(tab[, 1], tab[, 2])
summary(lm(tab[, 1] ~ tab[, 2]))
aov <- aov(diff ~ Ultra * Oil, data = oil.df)
qqnorm(aov$residuals)
qqline(aov$residuals)
plot(aov$residuals ~ aov$fitted.values)
detach()
```

The slope is far from 1, so we cannot say our model is additive. The normal plot and the residuals by fitted show concerning patterns as well.

## 5

### a

```{r}
ric.df <- read.csv("Ricci.csv", header =  T, stringsAsFactors = T)
attach(ric.df)
plot(Combine ~ Position)
detach()
```

There seems to be a slight different in mean scores.

### b

```{r}
attach(ric.df)
interaction.plot(Race, Position, Combine,
                 type = "b", pch = c(1, 4), col = c("red", "blue"))
detach()
```

The slopes are not parallel, so there seems to be interaction.

### c

```{r}
attach(ric.df)
tab <- tapply(Combine, list(Race, Position), mean)
tab
sendHelp <- data.frame(Position = c("Captain", "Captain", "Captain", "Lieutenant", "Lieutenant", "Lieutenant"), Race = c("B", "H", "W", "B", "H", "W"), Combine = c(63.7820, 68.5465, 74.1128, 63.71726, 63.62453, 71.84326))
library(lattice)
xyplot(sendHelp$Combine~sendHelp$Position,groups=sendHelp$Race,auto.key=list(space="right"),
       panel=function(x,y,...){
         panel.xyplot(x,y,...)
         panel.abline(h=mean(sendHelp$Combine),col="grey",lty=2,...)
       }
)
aov <- aov(Combine ~ Position * Race, data = ric.df)
qqnorm(aov$residuals)
qqline(aov$residuals)
plot(aov$residuals ~ aov$fitted.values)
detach(ric.df)
```

The residuals approximately follow the normal line and there is no obvious pattern in the resids vs fit. An additive model would also be appropriate.

### d

```{r}
lsdBW <- sqrt(72.6) * qt(0.975, 212) * sqrt(1/69 + 1/28)
lsdBH <- sqrt(72.6) * qt(0.975, 212) * sqrt(1/23 + 1/28)
lsdWH <- sqrt(72.6) * qt(0.975, 212) * sqrt(1/23 + 1/69)
lsdBW
lsdBH
lsdWH
attach(ric.df)
summary(aov)
mean <- tapply(Combine, Race, mean)
outer(mean, t(mean), "-")
```

Race is significant, but Position nor Position by Race are not.

Looking at LSD, as comparisons with White are all above the LSD, we can conclude that there was no type I error for differences between White and other races. However, this does not apply for the other races.


